---
title: 'Final Project: Tennis Match Prediction'
author: "Eric Browne, Oscar Pachon"
date: "5/24/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(GGally)
library(ggpubr)
library(lmtest)
library(MASS)
library(caret)
library(DescTools)
library(glmnet)
library(brglm2)
library(stringr)
library(enrichwith)
library(caTools)
library(ResourceSelection)
```

## Introduction

For our Final Project, we decided to create a Biased Reduced Generalized Linear model that predicts the outcome (win/lose) for a Tennis Match, in the form of a Logistic Regression.  The data that we have chosen has records across multiple tournaments for the best Male players, with explanatory variables correlating to attributes for each player coming into each match. 

## Setup / Cleaning

```{r}
# tennis<-read.csv("../Final_Project/Tennis_data.csv",header=FALSE)
# #tennis2<-read.csv("https://github.com/JeffSackmann/tennis_slam_pointbypoint/raw/master/2013-wimbledon-points.csv",header= FALSE)
# 
# 
# #view(tennis)
# #Need to add column names, and get rid of the first row
# #for (i in 1:54){
#  # names(tennis)[i] <- tennis[1,i]
# #}
# #Get rid of first row:
# #tennis <-tennis[-c(1),]
# #view(tennis$W4)
# 
# #Now treat the blank cells to a new data set, and see how many rows we lose:
# #Also, taking out the odds columns at the end, as we do not need them:
# #tennis1<-tennis1[-c(27)]
# ##  tennis1 will be out data set to be split into training, testing and validation sets
# view(tennis1)
# 
# #Now we must add columns to signify a higher rated player winning or not
# ##  VAriable: WNLSE
# ##    1: higher ranked player won
# ##    0: Lower ranked player won
# 
# #wnlse<-rep(1,46652)
# #tennis1<-cbind(tennis1,wnlse)
# #tennis1$WRank<-as.numeric(tennis1$WRank)
# #tennis1$LRank<-as.numeric(tennis1$LRank)
# #str(tennis1$LRank)
# #tennis1[tennis1==""]<-NA
# #Use an alias dataframe called tennis3
# #tennis3<-tennis1
# 
# #tennis3<-tennis3[!is.na(tennis3$WRank),]
# #tennis3<-tennis3[!is.na(tennis3$LRank),]
# 
# #is.na(tennis3$LRank)
# 
# 
# ## Loop Through the wnlse and fill the rows with the corresponding value
# 
# # for (i in 1:46358){
# #   tennis3$wnlse[i] = 0
# # }
# # for (i in 1:46358){
# #   if (tennis3$WRank[i] < tennis3$LRank[i]){
# #     tennis3$wnlse[i] = 1
# #   }
# # }
# 
# 
# length(unique(tennis3$Series)) #9 different types
# length(unique(tennis3$Court)) #2
# length(unique(tennis3$Surface)) #4
# length(unique(tennis3$Round)) # 9
# length(unique(tennis3$Best)) #6
# length(unique(tennis3$Winner)) # 876
# length(unique(tennis3$Loser)) # 1351
# length(unique(tennis3$Comment)) #12  #Not going to be used in regression
# 
# 
# 
# 
# Hard<-filter(tennis3, Surface == "Hard") #24272
# Carpet<-filter(tennis3, Surface == "Carpet") #1692
# Grass<-filter(tennis3, Surface == "Grass") #5183
# Clay<-filter(tennis3, Surface == "Clay") #15391
# 
# # we are going to throw out the observations for grass and carpet, since majority of games played are on Clay and Hard court
# 
# ## COME BACK HERE FOR TESTING ERRORS TO RESET DATASET
tennis_final<-filter(tennis3, Surface != "Carpet")
# tennis_final<-filter(tennis_final, Surface != "Grass")
# 
tennis_final$W1<-as.numeric(tennis_final$W1)
tennis_final$W2<-as.numeric(tennis_final$W2)
tennis_final$W3<-as.numeric(tennis_final$W3)
tennis_final$W4<-as.numeric(tennis_final$W4)
tennis_final$W5<-as.numeric(tennis_final$W5)
tennis_final$L1<-as.numeric(tennis_final$L1)
tennis_final$L2<-as.numeric(tennis_final$L2)
tennis_final$L3<-as.numeric(tennis_final$L3)
tennis_final$L4<-as.numeric(tennis_final$L4)
tennis_final$L5<-as.numeric(tennis_final$L5)
tennis_final$Wsets<-as.numeric(tennis_final$Wsets)
tennis_final$Lsets<-as.numeric(tennis_final$Lsets)
names(tennis_final)[9]<-"BestOf"
tennis_final$BestOf<-as.numeric(tennis_final$BestOf)
tennis_final<-filter(tennis_final, BestOf != "-1", BestOf != "0", BestOf !="1")


#Now using t.dat as the tennis data to model from
t.dat<-tennis_final

#Change the Series to be numeric
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("International Gold", "3", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("International Series", "6", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("International", "1", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("Grand Slam", "2", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("Masters Cup", "5", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("Masters 1000", "9", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("Masters", "4", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("ATP250", "7", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("ATP500", "8", x)}))

#Change Court to be Numeric
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("Outdoor", "1", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("Indoor", "2", x)}))
#Change Surface to be Numeric
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("Hard", "1", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("Clay", "2", x)}))
#Change Round to be Numeric
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("1st Round", "1", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("2nd Round", "2", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("Quarterfinals", "5", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("Semifinals", "6", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("The Final", "7", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("3rd Round", "3", x)}))
t.dat <- data.frame(lapply(t.dat, function(x) {
                  gsub("4th Round", "4", x)}))
t.dat<-t.dat[!grepl("Round Robin", t.dat$Round),]
t.dat<-t.dat[!grepl("0th Round", t.dat$Round),]

t.dat$Series<-as.numeric(t.dat$Series)
t.dat$Court<-as.numeric(t.dat$Court)
t.dat$Surface<-as.numeric(t.dat$Surface)
t.dat$Round<-as.numeric(t.dat$Round)
t.dat$W1<-as.numeric(t.dat$W1)
t.dat$W2<-as.numeric(t.dat$W2)
t.dat$W3<-as.numeric(t.dat$W3)
t.dat$W4<-as.numeric(t.dat$W4)
t.dat$W5<-as.numeric(t.dat$W5)
t.dat$L1<-as.numeric(t.dat$L1)
t.dat$L2<-as.numeric(t.dat$L2)
t.dat$L3<-as.numeric(t.dat$L3)
t.dat$L4<-as.numeric(t.dat$L4)
t.dat$L5<-as.numeric(t.dat$L5)
t.dat$Wsets<-as.numeric(t.dat$Wsets)
t.dat$Lsets<-as.numeric(t.dat$Lsets)
t.dat$WRank<-as.numeric(t.dat$WRank)
t.dat$LRank<-as.numeric(t.dat$LRank)
t.dat$wnlse<-as.numeric(t.dat$wnlse)
# str(t.dat)
# 
# unique(t.dat$Series) #9 different types
# unique(t.dat$Court) #2
# unique(t.dat$Surface) #2
# unique(t.dat$Round) #7
# 
# unique(t.dat$Round)
# 
# #Add in the difference in ranking
# t.dat2<-t.dat
# t.dat[c("WRank", "LRank")][is.na(t.dat[c("WRank", "LRank")])] <- 1500
# 
# max(unique(t.dat$WRank))
# for (i in 1:39347){
#    if(t.dat$WRank[i] > 1500){
#      t.dat$WRank[i] = 1500
#    }
#    if(t.dat$LRank[i] > 1500){
#      t.dat$LRank[i] = 1500
#    }
# 
#  }
# 
# 
# 
# 




#Save the file

save(testtennis2,file="tennisdata_small.Rdata")
save(t.dat,file="tennisdata_large.Rdata")

view(t.dat)

```

## Splitting the data into Training, Testing, and Validation

```{r}
# testtennis2<-testtennis
# testtennis2$Rankdiff<-abs(testtennis2$WRank-testtennis2$LRank)
#t.dat$Rankdiff<-abs(t.dat$WRank-t.dat$LRank)

logit_small<-glm(wnlse ~ factor(Surface)+WAce + LAce+ WRank + LRank+ W1 + W2 + L1 + L2, data=testtennis2, family = 'binomial')
summary(logit_small)


t.dat$Surface<-factor(t.dat$Surface)
t.dat$BestOf<-factor(t.dat$BestOf)
t.dat$Round<-factor(t.dat$Round)

logit_large_base<-glm(wnlse ~ factor(Surface)+factor(BestOf)+WAce + LAce+ WRank + LRank + W1 + W2+W3+W4+L1+L2+L3+L4 +Wsets+Lsets, data=t.dat, family = 'binomial')
summary(logit_large_base)
logit_large.2<-glm(wnlse ~ Surface+Round+ WRank + LRank+W2+L4 , data=t.dat2, family = 'binomial')
summary(logit_large.2)
length(which(t.dat2$Surface==2))
logit_large.3<-glm(wnlse ~ Surface, data=t.dat2, family = 'binomial')
summary(logit_large.3)


#large data: t.dat
non_upset_prop<-length(which(t.dat$wnlse == 1))/dim(t.dat)[1]
non_upset_prop #0.660
upset_prop<-length(which(t.dat$wnlse == 0))/dim(t.dat)[1]
upset_prop #0.340

unique(t.dat2$Rankdiff)

#small data: testtennis2
non_upset_prop.small<-length(which(testtennis2$wnlse == 1))/dim(testtennis2)[1]
non_upset_prop.small #0.593
upset_prop.small<-length(which(testtennis2$wnlse == 0))/dim(testtennis2)[1]
upset_prop.small # 0.407


#Filter out the ranks that are above 100:
t.dat100l<-filter(t.dat,WRank<100,LRank<100) #24794 Observations
#filter out the ranks that are above 200:
t.dat$rankdiff<-t.dat$Rankdiff
t.dat200l<-filter(t.dat,WRank<200,LRank<200)  #35136 observations

t.dat200l<-filter(t.dat200l,Comment=="Completed") #33686 Observations
t.dat200s<-t.dat200l
t.dat200s<-na.omit(t.dat200s) #1075 observations  #All 5 sets played

#Rank difference being greater than 100
t.dat_rankdiff200<-filter(t.dat,Rankdiff > 200)
t.dat_rankdiff5<-filter(t.dat,Rankdiff > 5)
t.dat_rankdiff100<-filter(t.dat,Rankdiff > 100)
t.dat_rankdiff500<-filter(t.dat,Rankdiff > 500)
#When rank difference is greater than 200, upsets occur 21% of the time
length(which(t.dat_rankdiff200$wnlse==0))/dim(t.dat_rankdiff200)[1] #0.21
#When rank difference is greater than 5, upsets occur 33% of the time
length(which(t.dat_rankdiff5$wnlse==0))/dim(t.dat_rankdiff5)[1] #0.33
#When rank difference is greater than 100, upsets occur 24% of the time
length(which(t.dat_rankdiff100$wnlse==0))/dim(t.dat_rankdiff100)[1] #0.24
#When rank difference is greater than 500, upsets occur 15% of the time
length(which(t.dat_rankdiff500$wnlse==0))/dim(t.dat_rankdiff500)[1] #0.15

##Now filter for the set counts on each BestOf Variable:
#Top 100
t.dat100b3<-filter(t.dat100l,BestOf==3)  #20605 Obersvations
length(which(t.dat100b3$wnlse==0))/20605  #36.6% upset
t.dat100b5<-filter(t.dat100l,BestOf==5)  #4189 Observations
length(which(t.dat100b5$wnlse==0))/4189  #27% upset

t.dat100b3.2<-filter(t.dat100b3, Wsets==2,Lsets==0) #12847 Observations
up1<-round(length(which(t.dat100b3.2$wnlse==0))/12847,3)  #33.7% upset
t.dat100b3.3<-filter(t.dat100b3, Wsets==2,Lsets==1) #6972 Observations
up2<-round(length(which(t.dat100b3.3$wnlse==0))/6972,3)  #40.99% upset

t.dat100b5.3<-filter(t.dat100b5, Wsets==3,Lsets==0)  #1989 Observations
up3<-round(length(which(t.dat100b5.3$wnlse==0))/1989,3) #19.5% upset
t.dat100b5.4<-filter(t.dat100b5, Wsets==3,Lsets==1)  #1239 Observations
up4<-round(length(which(t.dat100b5.4$wnlse==0))/1239,3) #29.9% upset
t.dat100b5.5<-filter(t.dat100b5, Wsets==3,Lsets==2)  #755 Observations
up5<-round(length(which(t.dat100b5.5$wnlse==0))/755,3) #42.4% upset

#Top 200
t.dat200b3<-filter(t.dat200l,BestOf==3)  #28003 Obersvations
t.dat200b5<-filter(t.dat200l,BestOf==5)  #5683 Observations

t.dat200b3.2<-filter(t.dat200b3, Wsets==2,Lsets==0) #18225 Observations
t.dat200b3.3<-filter(t.dat200b3, Wsets==2,Lsets==1) #9775 Observations

t.dat200b5.3<-filter(t.dat200b5, Wsets==3,Lsets==0)  #2862 Observations
t.dat200b5.4<-filter(t.dat200b5, Wsets==3,Lsets==1)  #1740 Observations
t.dat200b5.5<-filter(t.dat200b5, Wsets==3,Lsets==2)  #1074 Observations

#table of 100 cut bof 3, and 5
upset_percent100<-matrix(c(up1,"--",up2,up3,"--",up4,"--",up5),ncol=2,byrow=TRUE)
colnames(upset_percent100)<-c("Best of 3","Best of 5")
rownames(upset_percent100)<-c("2 Sets","3 Sets","4 Sets", "5 Sets")
upset_percent100<-as.table(upset_percent100)
upset_percent100


#Rankdiff on upsets
rd1<-round(mean(upset100.3.2$Rankdiff),3)
rd2<-round(mean(upset100.3.3$Rankdiff),3)
rd3<-round(mean(upset100.5.3$Rankdiff),3)
rd4<-round(mean(upset100.5.4$Rankdiff),3)
rd5<-round(mean(upset100.5.5$Rankdiff),3)
avgrankdiff_upset<-matrix(c(rd1,"---",rd2,rd3,"---",rd4,"---",rd5),ncol=2,byrow = TRUE)
colnames(avgrankdiff_upset)<-c("Best of 3", "Best of 5")
rownames(avgrankdiff_upset)<-c("2 Sets","3 Sets","4 Sets", "5 Sets")
avgrankdiff_upset100<-as.table(avgrankdiff_upset)
avgrankdiff_upset100

top25<-filter(t.dat100l,WRank<25,LRank<25) #2870 observations
top25.b3<-filter(top25,BestOf==3)
top25.b5<-filter(top25,BestOf==5)
top25.b3.2<-filter(top25,Wsets==2,Lsets==0)
upp1<-round(length(which(top25.b3.2$wnlse==0))/1369,3)
top25.b3.3<-filter(top25,Wsets==2,Lsets==1)
upp2<-round(length(which(top25.b3.3$wnlse==0))/780,3)
top25.b5.3<-filter(top25,Wsets==3,Lsets==0)
upp3<-round(length(which(top25.b5.3$wnlse==0))/299,3)
top25.b5.4<-filter(top25,Wsets==3,Lsets==1)
upp4<-round(length(which(top25.b5.4$wnlse==0))/200,3)
top25.b5.5<-filter(top25,Wsets==3,Lsets==2)
upp5<-round(length(which(top25.b5.5$wnlse==0))/119,3)

upset_percent25<-matrix(c(upp1,"--",upp2,upp3,"--",upp4,"--",upp5),ncol=2,byrow=TRUE)
colnames(upset_percent25)<-c("Best of 3","Best of 5")
rownames(upset_percent25)<-c("2 Sets","3 Sets","4 Sets", "5 Sets")
upset_percent25<-as.table(upset_percent25)
upset_percent25

rrd1<-round(mean(upset25.3.2$Rankdiff),3)
rrd2<-round(mean(upset25.3.3$Rankdiff),3)
rrd3<-round(mean(upset25.5.3$Rankdiff),3)
rrd4<-round(mean(upset25.5.4$Rankdiff),3)
rrd5<-round(mean(upset25.5.5$Rankdiff),3)
avgrankdiff_upset25<-matrix(c(rrd1,"---",rrd2,rrd3,"---",rrd4,"---",rrd5),ncol=2,byrow = TRUE)
colnames(avgrankdiff_upset25)<-c("Best of 3", "Best of 5")
rownames(avgrankdiff_upset25)<-c("2 Sets","3 Sets","4 Sets", "5 Sets")
avgrankdiff_upset25<-as.table(avgrankdiff_upset25)
avgrankdiff_upset25




```



## Charts Section:
```{r}
### RANK DIFFERENCES // RANKS ###
#Initial:
rankdifference1<-ggplot(data=t.dat)+geom_histogram(aes(x=Rankdiff,y=..count..)) +ggtitle("Initial Rank Differences") +labs(x="Difference in Ranking between Players",y="Count")
rankdifference1

#After cut of players below rank 200:
rankdifference2<-ggplot(data=t.dat200l)+geom_histogram(aes(x=Rankdiff,y=..count..)) + ggtitle("Players Below Rank: 200")+labs(x="Difference in Ranking between players",y="Count")
rankdifference2

##Amount of upsets on the different BestOf data sets:   Similar behavior in sub 200 and sub 100 ranks
#After cut 100, 2/3 sets played
upset100.3.2<-filter(t.dat100b3.2,wnlse==0) #4336 obs
dim(upset100.3.2)[1]
#After cut 100, 3/3 sets played
upset100.3.3<-filter(t.dat100b3.3,wnlse==0) #2858 obs
dim(upset100.3.3)[1]
#After cut 100, 3/5 sets played
upset100.5.3<-filter(t.dat100b5.3,wnlse==0) #387 obs
dim(upset100.5.3)[1]
#After cut 100, 4/5 sets played
upset100.5.4<-filter(t.dat100b5.4,wnlse==0) #370 obs
dim(upset100.5.4)[1]
#After cut 100, but went all 5 sets:
upset100.5.5<-filter(t.dat100b5.5,wnlse==0) #320 obs
dim(upset100.5.5)[1]
#After cut 200, 2/3 sets played
upset200.3.2<-filter(t.dat200b3.2,wnlse==0) #6106 obs
dim(upset200.3.2)[1]
#After cut 100, 3/3 sets played
upset200.3.3<-filter(t.dat200b3.3,wnlse==0) #3959 obs
dim(upset200.3.3)[1]
#After cut 200, 3/5 sets played
upset200.5.3<-filter(t.dat200b5.3,wnlse==0) #554 obs
dim(upset200.5.3)[1]
#After cut 200, 4/5 sets played
upset200.5.4<-filter(t.dat200b5.4,wnlse==0) #516 obs
dim(upset200.5.4)[1]
#After cut 200, but went all 5 sets:
upset200.5.5<-filter(t.dat200b5.5,wnlse==0) #449 obs
dim(upset200.5.5)[1]

#After cut 25, 2/3 sets played
upset25.3.2<-filter(top25.b3.2, wnlse==0) #432 obs
dim(upset25.3.2)[1]
#After cut 25, 3/3 sets played
upset25.3.3<-filter(top25.b3.3,wnlse==0) #332 obs
dim(upset25.3.3)[1]
#After cut 25, 3/5 sets played
upset25.5.3<-filter(top25.b5.3,wnlse==0) #58 obs
dim(upset25.5.3)[1]
#After cut 25, 4/5 sets played
upset25.5.4<-filter(top25.b5.4,wnlse==0) #70 obs
dim(upset25.5.4)[1]
#After cut 25, but went all 5 sets:
upset25.5.5<-filter(top25.b5.5,wnlse==0) #46 obs
dim(upset25.5.5)[1]


rankdifference3<-ggplot(data=t.dat200s)+geom_histogram(aes(x=Rankdiff,y=..count..)) + ggtitle("Players Below Rank: 200, All 5 Sets Played")+labs(x="Difference in Ranking between players",y="Count")
rankdifference3


#Take average rank difference when upset occurs:
upset_only<-filter(t.dat, wnlse==0)
upset_only100<-filter(t.dat100l,wnlse==0)
upset_only200<-filter(t.dat200l,wnlse==0)
upset_only25<-filter(top25,wnlse==0)
upset_only100b3<-filter(t.dat100b3,wnlse==0)
upset_only100b5<-filter(t.dat100b5,wnlse==0)
mean(upset_only$Rankdiff)
mean(upset_only100$Rankdiff)
rankdifference4<-ggplot(data=upset_only)+geom_histogram(aes(x=Rankdiff, y=..count..))+ggtitle("Rank Difference when an upset occurs")+labs(x="Difference in Rank",y='Count')
rankdifference5<-ggplot(data=upset_only100)+geom_histogram(aes(x=Rankdiff, y=..count..))+ggtitle("Rank Difference when an Upset Occurs, post 100 cut")+labs(x="Difference in Rank",y='Count')
rankdifference4 #mean 55
rankdifference5 #mean 25

rankdifference6<-ggplot(data=upset_only100b3)+geom_histogram(aes(x=Rankdiff, y=..count..))+ggtitle("Rank Difference when an upset occurs, BOF 3/3 Sets: Post 100 Cut")+labs(x="Difference in Rank",y='Count')
rankdifference7<-ggplot(data=upset_only100b5)+geom_histogram(aes(x=Rankdiff, y=..count..))+ggtitle("Rank Difference when an upset occurs, BOF 5/5 Sets: Post 100 Cut")+labs(x="Difference in Rank",y='Count')





#Original un-cut data hist of ranks
rank1<-ggplot(data=t.dat)+geom_histogram(aes(x=WRank,y=..count..))+ggtitle("Winner's Rankings without Rank Cutoff")+labs(x="Winner's Ranking")
rank1
#Post 200 cut
rank2<-ggplot(data=t.dat200l)+geom_histogram(aes(x=WRank,y=..count..))+ggtitle("Winner's Rankings, after 200 Cutoff")+labs(x="Winner's Ranking")
rank2
#Post 100 cut
rank3<-ggplot(data=t.dat100l)+geom_histogram(aes(x=WRank,y=..count..))+ggtitle("Winner's Rankings, after 100 Cutoff")+labs(x="Winner's Ranking")
rank3

#Upsets only rankings:
rank4<-ggplot(data=upset_only)+geom_histogram(aes(x=WRank,y=..count..))+ggtitle("Winners Rankings of Upsets")+labs(x="Winner's Rank during Upset")
rank4

rank5<-ggplot(data=upset_only100)+geom_histogram(aes(x=WRank,y=..count..))+ggtitle("Winners Rankings of Upsets: Top 100")+labs(x="Winner's Rank during Upset")
rank5

rank6<-ggplot(data=upset_only200)+geom_histogram(aes(x=WRank,y=..count..))+ggtitle("Winners Rankings of Upsets: Top 200")+labs(x="Winner's Rank during Upset")
rank6

rank7<-ggplot(data=upset_only25)+geom_histogram(aes(x=WRank,y=..count..))+ggtitle("Winners Rankings of Upsets: Top 25")+labs(x="Winner's Rank during Upset")
rank7





### PLAYER COUNT ###
##Player count
length(unique(t.dat200l$Winner))  #615 different winners
length(unique(t.dat200l$Loser))  #765 different losers
length(unique(which(t.dat200l$Winner=="Federer R.")))
length(which(t.dat200l$Loser=="Federer R."))
length(which(t.dat200l$Winner=="Federer R."))
barplot(table(t.dat200l$Winner))
barplot(table(t.dat200l$Loser))
playercount<-ggplot(data=t.dat200l)+geom_point(aes(x=Winner,y=..count..))
playercount
```

## Models Finally:
```{r}
#Replace all NA values in W1-5, L1-5 with 0s for regression
x[c("a", "b")][is.na(x[c("a", "b")])] <- 0
t.dat[c("W1","W2","W3","W4","W5","L1","L2","L3","L4","L5")][is.na(t.dat[c("W1","W2","W3","W4","W5","L1","L2","L3","L4","L5")])]<-0
top25[c("W1","W2","W3","W4","W5","L1","L2","L3","L4","L5")][is.na(top25[c("W1","W2","W3","W4","W5","L1","L2","L3","L4","L5")])]<-0
t.dat100l[c("W1","W2","W3","W4","W5","L1","L2","L3","L4","L5")][is.na(t.dat100l[c("W1","W2","W3","W4","W5","L1","L2","L3","L4","L5")])]<-0
t.dat100b3<-t.dat100b3[-c("W4","W5","L4","L5")]
t.dat_train$wnlse<-as.numeric(t.dat_train$wnlse)

logit.1100<-glm(wnlse ~ factor(Surface)+factor(BestOf)+WAce + LAce+ WRank + LRank + W1 + W2+W3+W4+L1+L2+L3+L4 +Wsets+Lsets, data=t.dat100l, family = 'binomial')
logit_all<-glm(wnlse ~ factor(Surface)+factor(BestOf)+WAce + LAce+ WRank + LRank + W1 + W2+W3+W4+L1+L2+L3+L4 +Wsets+Lsets, data=t.dat_train, method ='brglmFit')
logit.100<-glm(wnlse ~ factor(Surface)+factor(BestOf)+WAce + LAce+ WRank + LRank + W1 + W2+W3+W4+L1+L2+L3+L4 +Wsets+Lsets, data=t.dat100l_train, method ='brglmFit')
logit.25<-glm(wnlse ~ factor(Surface)+WAce + LAce+ WRank + LRank + W1 + W2+W3+W4+L1+L2+L3+L4 +Wsets+Lsets, data=top25_train, method ='brglmFit')

logit_allrd<-

#Second Round after eliminating variables:
logitallv2<-glm(wnlse ~ WRank + LRank + W1 + W2+W3+L1+L2+L3 +Wsets, data=t.dat_train, method ='brglmFit')
logit100v2<-glm(wnlse ~ WRank + LRank, data=t.dat100l_train, method = 'brglmFit')
logit25v2<-glm(wnlse ~ WRank + LRank+W1, data=top25_train, method = 'brglmFit')

logit.100bof3.3<-glm(wnlse ~ factor(Surface)+WAce + LAce+ WRank + LRank + W1 + W2+W3+L1+L2+L3 +Wsets+Lsets, data=t.dat100b3.3, method ='brglmFit')
logit.100bof3.3v2<-glm(wnlse ~WAce+ WRank + LRank+L3, data=t.dat100b3.3, method ='brglmFit')

logit.100bof5.5<-glm(wnlse ~ LAce+ WRank + LRank + W1 + W2+W3+W4+W5+L1+L2+L3+L4+L5 +Wsets+Lsets, data=t.dat100b5.5, method ='brglmFit')
logit.100bof5.5v2<-glm(wnlse ~ LAce+ WRank + LRank, data=t.dat100b5.5, method ='brglmFit')

#Anova/ LR test #error
anova_all<-lrtest(logit_all,logitallv2)
anova_100<-anova(logit.100,logit100v2,test="F")
anova_25<-anova(logit_all,logitallv2,test="F")

#Hosmer Test #error
hoslem.test(logit100v2$wnlse, fitted(logit25v2), g = 10)

summary(logit.100)
summary(logit.25)
summary(logit100v2)
summary(logit25v2)
summary(logit_all)
summary(logitallv2)
summary(logit.100bof3.3)
summary(logit.100bof3.3v2)
summary(logit.100bof5.5)
summary(logit.100bof5.5v2)




##Split the data into training, testing and validating:

#All
t.dat$wnlse<-as.factor(t.dat$wnlse)
set.seed(42069)
trainIndex<-createDataPartition(t.dat$wnlse,p=0.7,list=FALSE,times=1)

t.dat.train<-t.dat[trainIndex,]
t.dat.test<-t.dat[-trainIndex,]
#All
n=nrow(t.dat)
t.dat$tvt<-sample(rep(0:2,c(round(n*.5),round(n*.25),n-2*round(n*.25))),n)
t.dat_test <- t.dat %>% slice(which(t.dat$tvt==2))
t.dat_train <- t.dat %>% slice(which(t.dat$tvt ==0))
t.dat_val <- t.dat %>% slice(which(t.dat$tvt==1))



#Top 100 Rank:
nn=nrow(t.dat100l)
t.dat100l$tvt<-sample(rep(0:2,c(round(nn*.5),round(nn*.25),nn-2*round(nn*.25))),nn)
t.dat100l_test <- t.dat100l %>% slice(which(t.dat100l$tvt==2))
t.dat100l_train <- t.dat100l %>% slice(which(t.dat100l$tvt ==0))
t.dat100l_val <- t.dat100l %>% slice(which(t.dat100l$tvt==1))


#Top 25 Rank:
nn=nrow(top25)
top25$tvt<-sample(rep(0:2,c(round(nn*.5),round(nn*.25),nn-2*round(nn*.25))),nn)
top25_test<- top25 %>% slice(which(top25$tvt==2))
top25_train <- top25 %>% slice(which(top25$tvt ==0))
top25_val <- top25 %>% slice(which(top25$tvt==1))

#Confusion Matrices:
##All
p<-predict(logitallv2,t.dat_test,type="response")
prediction<-p>=.5
confusion_mat.all <- table(t.dat_test$wnlse,prediction)
confusion_mat.all
all_Acc<-(confusion_mat.all[1]+confusion_mat.all[4])/sum(confusion_mat.all) #0.667 accuracy
dim(t.dat_test)[1] #15732


##Top 100
p<-predict(logit100v2,t.dat100l_test,type="response")
prediction<-p>=.5
confusion_mat.100 <- table(t.dat100l_test$wnlse,prediction)
confusion_mat.100
top100_Acc<-(confusion_mat.100[1]+confusion_mat.100[4])/sum(confusion_mat.100) #0.9683998 accuracy
dim(t.dat100l_test)[1] #9905



##Top 25 Accuracy
p<-predict(logit25v2,top25_test,type="response")
prediction<-p>=.5
confusion_mat.25 <- table(top25_test$wnlse,prediction)
confusion_mat.25
top25_Acc<-(confusion_mat.25[1]+confusion_mat.25[4])/sum(confusion_mat.25)#0.971478 accuracy
dim(top25_test)[1] #1157

##Can do more on confusion matrices, but dont really need to, but can if we want: top100bof3,top100bof5


```

## Load in Data sets for Each R Session:
```{r}
#Load data sets:
load("tennisdata_large.Rdata")  #t.dat
load("tennisdata_small.Rdata")  #testtennis2

save(t.dat100l,file="tennisdata_top100.Rdata")
save(t.dat,file="tennisdata_tdat.Rdata")
save(top25,file="tennisdata_top25.Rdata")




##Charts So Far:
rankdifference1
rankdifference2
rankdifference3
rankdifference4
rankdifference5
rankdifference6
rankdifference7

rank1
rank2
rank3
rank4
rank5
rank6
rank7


##Tables:
upset_percent100
upset_percent25
avgrankdiff_upset100
avgrankdiff_upset25
upset_percent

##Proportions:
#Upsets in All:
per1<-round(length(which(t.dat$wnlse==0))/dim(t.dat)[1],3) #33%upset
per2<-round(length(which(t.dat100l$wnlse==0))/dim(t.dat100l)[1],3) #35%upset
per3<-round(length(which(top25$wnlse==0))/dim(top25)[1],3)

upset_percent<-matrix(c(per1,per2,per3),ncol=1,byrow = TRUE)
colnames(upset_percent)<-c("Percent of Upsets")
rownames(upset_percent)<-c("All Ranks","Top 100 Ranks","Top 25 Ranks")
upset_percent<-as.table(upset_percent)
upset_percent

##Regression Models:
summary(logit.100)
summary(logit.25)
summary(logit100v2)
summary(logit25v2)
summary(logit_all)
summary(logitallv2)
summary(logit.100bof3.3)
summary(logit.100bof3.3v2)
summary(logit.100bof5.5)
summary(logit.100bof5.5v2)

confusion_mat.all
confusion_mat.25
confusion_mat.100
all_Acc
top100_Acc
top25_Acc

dim(t.dat)
```
